---
title: "final-story"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the janitor and tidyverse packages
```{r}
#install.packages("janitor")
library(janitor)
#install.packages("tidyverse")
library(tidyverse)
library(lubridate)
```


### Reporting Question
#### I noticed that a large amount of donations came from Kim Coley herself as in-kind donations. I'm curious about how well campaigns turn out depending on how much the candidate invests in themselves. Is there a correlation between how much money a candidate invests in their own campaign and how much money others invest in their campaign? Do candidates who invest a certain amount into their campaign have a higher chance of winning?

### Get campaign finance data
#### This data includes all 2018 quarterly reports for 2018 NC Senate candidates where available
```{r}

# Reads in candidate data
candidateData <- read_csv("../../python/senate2018/urls.csv")

receiptsAll <- list()

# Given a district number and last name, consolidates the quarterly receipts for a candidate
# Returns a dataframe of receipts from 2018
consolidateReceipts <- function(district, lastname) {
  # Creates an empty dataframe with column names
  rcpts_complete <- data.frame(matrix(vector(), 0, 19,
                dimnames=list(c(), c("Date", "Is Prior", "Name", "Street 1", "Street 2", "City", "State", "Full Zip", 
                                     "Country Name", "Outside US Postal Code", "Profession", "Employers Name", "Purpose", 
                                     "Receipt Type Desc", "Account Abbr", "Form Of Payment Desc", "Description", "Amount", 
                                     "Sum To Date"))),
                stringsAsFactors=F)
  
    # Constructs CSV paths from district and last name
    urls <- c(sprintf("../../python/data/senate2018/%s/%s/q1.csv", district, lastname),
              sprintf("../../python/data/senate2018/%s/%s/q2.csv", district, lastname),
              sprintf("../../python/data/senate2018/%s/%s/q3.csv", district, lastname),
              sprintf("../../python/data/senate2018/%s/%s/q4.csv", district, lastname))
    
    # Tries to open each quarterly receipt CSV and unions it to a running dataframe
    for(url in urls){
      # If the file doesn't exist, skip it
      if (file.exists(url)) {
        temp <- read_csv(url, 
          col_types = cols(
            `Account Abbr` = col_character(), 
              City = col_character(), 
            `Country Name` = col_character(), 
              Date = col_date(format = "%m/%d/%Y"), 
              Description = col_character(), 
            `Employers Name` = col_character(), 
              `Full Zip` = col_character(), 
            `Outside US Postal Code` = col_character(), 
              Profession = col_character(), 
            Purpose = col_character(), 
              State = col_character(), 
            `Street 1` = col_character(), 
              `Street 2` = col_character()),
          skip = 1)
        
        rcpts_complete <- union_all(rcpts_complete, temp)
      }
    }
    
    #Finally, we should rename the columns to remove spaces and generally promote brevity.
    names(rcpts_complete) <- c("date","prior","donor","street1","street2","city","state","zip","country","postal","profession","employer","purpose","type","account","payment_form","description","amount","sum_to_date")
    
    return(rcpts_complete)
}

#candidateData <- candidateData %>%
#  mutate(receipts2018 = list(consolidateReceipts(district, tolower(lastname))))

# For each candidate, consolidate their quarterly receipts
for (row in 1:nrow(candidateData)) {
    district <- candidateData[row, "district"]
    lastname <- candidateData[row, "lastname"] %>% tolower()
    
    rcpts_complete <- consolidateReceipts(district, lastname)

    receiptsAll[[row]] <- rcpts_complete
}

# Update the candidate data with their receipts as lists
candidateData$receipts2018 <- receiptsAll

```

### Calculate self funding
#### Searches through each receipt data frame for matches with the candidate's name
```{r}

# Given a dataframe of receipts, a last name, and a first name/aliases, calculates the amount of self funding a candidate received
# Returns the total amount received through self funding
calculateSelfFunding <- function(receipts, lastname, firstnames) {
  # Regex pattern that matches any instance where the candidate's first name/aliases precedes their last name, with anything in between and at the beginning and end of strings
  pattern <- sprintf(".*%s.*%s.*", paste(firstnames, collapse = "|"), lastname)
  
  # Filters by using grepping with the above pattern on the donors
  receipts <- receipts %>%
    filter(grepl(pattern, c(receipts2018.donor), ignore.case=TRUE))
  
  # Uncomment this print statement to check for discrepancy in resulting list of self-funded receipts
  # print(receipts)
  
  total <- sum(receipts$receipts2018.amount)
  return(total)
}

selfFundedAll <- c()

# For each candidate, calculate their self-funded total
for (row in 1:nrow(candidateData)) {
    district <- candidateData[row, "district"]
    lastname <- candidateData[row, "lastname"] %>% tolower()
    firstnames <- c(candidateData[row, "firstname"] %>% tolower())
    aliases <- candidateData[row, "aliases"]
    if (!(is.na(aliases) || aliases == "") && typeof(aliases) == "character") {
      firstnames <- append(firstnames, strsplit(candidateData[row, "aliases"], "/"))
    }
    receipts <- candidateData[row, "receipts2018"]
    receipts <- do.call(data.frame, receipts)
    
    selfFundedAmt <- calculateSelfFunding(receipts, lastname, firstnames)

    selfFundedAll <- append(selfFundedAll, selfFundedAmt)
}

# Update the candidate data with their self-funded totals
candidateData$selfFundedAmt <- selfFundedAll

```

### Calculate relative self funding compared to opponent
#### For candidate, we'll get the relative funding compared to their district opponent in the other party
```{r}

candidateDataNaN <- candidateData %>%
  group_by(district) %>%
  mutate(selfFundedTotal = sum(selfFundedAmt), selfFundedRelative = selfFundedAmt / (selfFundedTotal)) 

candidateDataClean <- candidateDataNaN %>%
  mutate(selfFundedRelative = if(!is.finite(selfFundedRelative)) 0.5 else selfFundedRelative)

```

### Calculate the percentage of campaigns won by the highest self funder
```{r}
nrow(candidateDataClean %>%
  filter(win == 1, selfFundedRelative > 0.5))

nrow(candidateDataClean %>%
  filter(win == 1, selfFundedRelative < 0.5))
```


### Plot relative self funding vs margin of victory for winners
```{r}

candidateDataClean %>%
  filter(win == 1) %>%
  ggplot(mapping = aes(x = selfFundedRelative, y = margin, size = selfFundedAmt)) + 
  geom_point() +
  ggtitle("Self Funding vs Margin of Victory") +
  xlab("Relative Self Funding") + 
  ylab("Margin of Victory (percentage)") +
  labs(
    size = "Self Funding ($)"
  ) +
  geom_smooth()

candidateDataClean %>%
  filter(win == 1, selfFundedTotal > 1000) %>%
  ggplot(mapping = aes(x = selfFundedRelative, y = margin, size = selfFundedAmt)) + 
  geom_point() + 
  ggtitle("Self Funding vs Margin of Victory") +
  xlab("Relative Self Funding") + 
  ylab("Margin of Victory (percentage)") +
  labs(
    size = "Self Funding ($)"
  ) +
  geom_smooth()

```

### Remove all races where neither party self funded
```{r}

candidateDataNaN %>%
  filter(win == 1, is.finite(selfFundedRelative)) %>%
  ggplot(mapping = aes(x = selfFundedRelative, y = margin, size = selfFundedAmt)) + 
  geom_point() + 
  ggtitle("Self Funding vs Margin of Victory") +
  xlab("Relative Self Funding") + 
  ylab("Margin of Victory (percentage)") +
  labs(
    size = "Self Funding ($)"
  ) +
  geom_smooth()

cor(candidateDataClean$selfFundedRelative, candidateDataClean$margin)

```

### Gather close funding data
#### Get data donated from family and family businesses
```{r}

candidateDataClose <- candidateData

# Given a dataframe of receipts, a last name, and a first name/aliases, calculates the amount of self funding a candidate received
# Returns the total amount received through self funding
calculateCloseFunding <- function(receipts, lastname) {
  # Regex pattern that matches any instance where the candidate's first name/aliases precedes their last name, with anything in between and at the beginning and end of strings
  pattern <- sprintf(".*%s.*", lastname)
  
  # Filters by using grepping with the above pattern on the donors
  receipts <- receipts[ grepl(pattern, receipts$receipts2018.donor, ignore.case=TRUE) | grepl(pattern, receipts$receipts2018.employer, ignore.case=TRUE), ]
  
  # Uncomment this print statement to check for discrepancy in resulting list of self-funded receipts
  # print(receipts)
  
  total <- sum(receipts$receipts2018.amount)
  return(total)
}

closeFundedAll <- c()

# For each candidate, calculate their self-funded total
for (row in 1:nrow(candidateData)) {
    district <- candidateData[row, "district"]
    lastname <- candidateData[row, "lastname"] %>% tolower()
    receipts <- candidateData[row, "receipts2018"]
    receipts <- do.call(data.frame, receipts)
    
    closeFundedAmt <- calculateCloseFunding(receipts, lastname)

    closeFundedAll <- append(closeFundedAll, closeFundedAmt)
}

# Update the candidate data with their self-funded totals
candidateDataClose$closeFundedAmt <- closeFundedAll

candidateDataClose$closeFundedAmt

```

### Calculate relative close funding compared to opponent
```{r}

candidateDataCloseNaN <- candidateDataClose %>%
  group_by(district) %>%
  mutate(closeFundedTotal = sum(closeFundedAmt), closeFundedRelative = closeFundedAmt / (closeFundedTotal)) 

candidateDataCloseClean <- candidateDataCloseNaN %>%
  mutate(closeFundedRelative = if(!is.finite(closeFundedRelative)) 0.5 else closeFundedRelative)

```

### Calculate the percentage of campaigns won by the highest close funder
```{r}
nrow(candidateDataCloseClean %>%
  filter(win == 1, closeFundedRelative > 0.5))

nrow(candidateDataCloseClean %>%
  filter(win == 1, closeFundedRelative < 0.5))
```


### Plot relative funding vs margin of victory for winners
```{r}

candidateDataCloseClean %>%
  filter(win == 1) %>%
  ggplot(mapping = aes(x = closeFundedRelative, y = margin, size = closeFundedAmt)) + 
  geom_point() + 
  ggtitle("Close Funding vs Margin of Victory") +
  xlab("Relative Close Funding") + 
  ylab("Margin of Victory (percentage)") +
  labs(
    size = "Close Funding ($)"
  ) +
  geom_smooth()

candidateDataCloseClean %>%
  filter(win == 1, closeFundedTotal > 1000) %>%
  ggplot(mapping = aes(x = closeFundedRelative, y = margin, size = closeFundedAmt)) + 
  geom_point() + 
  ggtitle("Close Funding vs Margin of Victory") +
  xlab("Relative Close Funding") + 
  ylab("Margin of Victory (percentage)") +
  labs(
    size = "Close Funding ($)"
  ) +
  geom_smooth()

```

### Remove all races where neither party is close funded
```{r}

candidateDataCloseNaN %>%
  filter(win == 1, is.finite(closeFundedRelative)) %>%
  ggplot(mapping = aes(x = closeFundedRelative, y = margin, size = closeFundedAmt)) + 
  geom_point() + 
  ggtitle("Close Funding vs Margin of Victory") +
  xlab("Relative Close Funding") + 
  ylab("Margin of Victory (percentage)") +
  labs(
    size = "Close Funding ($)"
  ) +
  geom_smooth()

cor(candidateDataCloseClean$closeFundedRelative, candidateDataCloseClean$margin)

```

### Calculate self funding amounts before the primary
```{r}

candidateDataEarly <- candidateData

# Given a dataframe of receipts, a last name, and a first name/aliases, calculates the amount of early self funding a candidate received
# Returns the total amount received through early self funding
calculateEarlySelfFunding <- function(receipts, lastname, firstnames) {
  # Regex pattern that matches any instance where the candidate's first name/aliases precedes their last name, with anything in between and at the beginning and end of strings
  pattern <- sprintf(".*%s.*%s.*", paste(firstnames, collapse = "|"), lastname)
  
  # Filters by using grepping with the above pattern on the donors
  receipts <- receipts %>%
    filter(grepl(pattern, c(receipts2018.donor), ignore.case=TRUE), ymd(receipts2018.date) < mdy("5/8/2018"))
  
  # Uncomment this print statement to check for discrepancy in resulting list of early self-funded receipts
  # print(receipts)
  
  total <- sum(receipts$receipts2018.amount)
  return(total)
}

earlySelfFundedAll <- c()

# For each candidate, calculate their early self-funded total
for (row in 1:nrow(candidateData)) {
    district <- candidateData[row, "district"]
    lastname <- candidateData[row, "lastname"] %>% tolower()
    firstnames <- c(candidateData[row, "firstname"] %>% tolower())
    # <- append(strsplit(candidateData[row, "aliases"], "/"), candidateData[row, "firstname"])
    receipts <- candidateData[row, "receipts2018"]
    receipts <- do.call(data.frame, receipts)
    
    earlySelfFundedAmt <- calculateEarlySelfFunding(receipts, lastname, firstnames)

    earlySelfFundedAll <- append(earlySelfFundedAll, earlySelfFundedAmt)
}

# Update the candidate data with their self-funded totals
candidateDataEarly$earlySelfFundedAmt <- earlySelfFundedAll

```

### Calculate relative early self funding compared to opponent
#### For candidate, we'll get the relative funding compared to their district opponent in the other party
```{r}

candidateDataEarlyNaN <- candidateDataEarly %>%
  group_by(district) %>%
  mutate(earlySelfFundedTotal = sum(earlySelfFundedAmt), earlySelfFundedRelative = earlySelfFundedAmt / (earlySelfFundedTotal)) 

candidateDataEarlyClean <- candidateDataEarlyNaN %>%
  mutate(earlySelfFundedRelative = if(!is.finite(earlySelfFundedRelative)) 0.5 else earlySelfFundedRelative)

for (row in 1:nrow(candidateDataEarlyNaN)) {
  receipts <- candidateDataEarlyNaN[row, "receipts2018"]
  receipts <- do.call(data.frame, receipts)
  candidateDataEarlyNaN[row, "totalAmount"] = sum(receipts$receipts2018.amount)
}

for (row in 1:nrow(candidateDataEarlyClean)) {
  receipts <- candidateDataEarlyClean[row, "receipts2018"]
  receipts <- do.call(data.frame, receipts)
  candidateDataEarlyClean[row, "totalAmount"] = sum(receipts$receipts2018.amount)
}

candidateDataEarlyNaN <- candidateDataEarlyNaN %>%
  group_by(district) %>%
  mutate(combinedTotal = sum(totalAmount), totalRelative = totalAmount / (combinedTotal)) 

candidateDataEarlyClean <- candidateDataEarlyClean %>%
  group_by(district) %>%
  mutate(combinedTotal = sum(totalAmount), totalRelative = totalAmount / (combinedTotal)) %>%
  mutate(totalRelative = if(!is.finite(totalRelative)) 0.5 else totalRelative)

```

### Plot early self funding against total amount raised
```{r}

candidateDataEarlyClean %>%
  filter(win == 1) %>%
  ggplot(mapping = aes(x = earlySelfFundedRelative, y = totalRelative, size = earlySelfFundedAmt)) + 
  geom_point() + 
  ggtitle("Early Self Funding vs Total Amount Raised") +
  xlab("Relative Early Self Funding") + 
  ylab("Relative Total Raised") +
  labs(
    size = "Early Self Funding ($)"
  ) +
  geom_smooth()

candidateDataEarlyNaN %>%
  filter(win == 1, is.finite(earlySelfFundedRelative)) %>%
  ggplot(mapping = aes(x = earlySelfFundedRelative, y = totalRelative, size = earlySelfFundedAmt)) + 
  geom_point() + 
  ggtitle("Early Self Funding vs Total Amount Raised") +
  xlab("Relative Early Self Funding") + 
  ylab("Relative Total Raised") +
  labs(
    size = "Early Self Funding ($)"
  ) +
  geom_smooth()

```

### Compare early self funding to margin of victory
```{r}
candidateDataEarlyClean %>%
  #filter(earlySelfFundedAmt > 0) %>%
  filter(win == 1) %>%
  ggplot(mapping = aes(x = earlySelfFundedRelative, y = margin)) + 
  geom_point(size=3) + 
  geom_smooth()
```

### Calculate close funding amounts before the primary
```{r}

candidateDataEarlyClose <- candidateData

# Given a dataframe of receipts, a last name, and a first name/aliases, calculates the amount of early self funding a candidate received
# Returns the total amount received through early close funding
calculateEarlyCloseFunding <- function(receipts, lastname, firstnames) {
  # Regex pattern that matches any instance where the candidate's first name/aliases precedes their last name, with anything in between and at the beginning and end of strings
  pattern <- sprintf(".*%s.*", lastname)
  
  # Filters by using grepping with the above pattern on the donors
  receipts <- receipts[ grepl(pattern, receipts$receipts2018.donor, ignore.case=TRUE) | grepl(pattern, receipts$receipts2018.employer, ignore.case=TRUE), ]
  receipts <- receipts %>%
    filter(ymd(receipts2018.date) < mdy("5/8/2018"))
  
  # Uncomment this print statement to check for discrepancy in resulting list of early close-funded receipts
  # print(receipts)
  
  total <- sum(receipts$receipts2018.amount)
  return(total)
}

earlyCloseFundedAll <- c()

# For each candidate, calculate their early close-funded total
for (row in 1:nrow(candidateData)) {
    district <- candidateData[row, "district"]
    lastname <- candidateData[row, "lastname"] %>% tolower()
    firstnames <- c(candidateData[row, "firstname"] %>% tolower())
    # <- append(strsplit(candidateData[row, "aliases"], "/"), candidateData[row, "firstname"])
    receipts <- candidateData[row, "receipts2018"]
    receipts <- do.call(data.frame, receipts)
    
    earlyCloseFundedAmt <- calculateEarlyCloseFunding(receipts, lastname, firstnames)

    earlyCloseFundedAll <- append(earlyCloseFundedAll, earlyCloseFundedAmt)
}

# Update the candidate data with their close-funded totals
candidateDataEarlyClose$earlyCloseFundedAmt <- earlyCloseFundedAll

```


### Calculate relative early self funding compared to opponent
#### For candidate, we'll get the relative funding compared to their district opponent in the other party
```{r}

candidateDataEarlyCloseNaN <- candidateDataEarlyClose %>%
  group_by(district) %>%
  mutate(earlyCloseFundedTotal = sum(earlyCloseFundedAmt), earlyCloseFundedRelative = earlyCloseFundedAmt / (earlyCloseFundedTotal)) 

candidateDataEarlyCloseClean <- candidateDataEarlyCloseNaN %>%
  mutate(earlyCloseFundedRelative = if(!is.finite(earlyCloseFundedRelative)) 0.5 else earlyCloseFundedRelative)

for (row in 1:nrow(candidateDataEarlyCloseNaN)) {
  receipts <- candidateDataEarlyCloseNaN[row, "receipts2018"]
  receipts <- do.call(data.frame, receipts)
  candidateDataEarlyCloseNaN[row, "totalAmount"] = sum(receipts$receipts2018.amount)
}

for (row in 1:nrow(candidateDataEarlyCloseClean)) {
  receipts <- candidateDataEarlyCloseClean[row, "receipts2018"]
  receipts <- do.call(data.frame, receipts)
  candidateDataEarlyCloseClean[row, "totalAmount"] = sum(receipts$receipts2018.amount)
}

candidateDataEarlyCloseNaN <- candidateDataEarlyCloseNaN %>%
  group_by(district) %>%
  mutate(combinedTotal = sum(totalAmount), totalRelative = totalAmount / (combinedTotal)) 

candidateDataEarlyCloseClean <- candidateDataEarlyCloseClean %>%
  group_by(district) %>%
  mutate(combinedTotal = sum(totalAmount), totalRelative = totalAmount / (combinedTotal)) %>%
  mutate(totalRelative = if(!is.finite(totalRelative)) 0.5 else totalRelative)

```

### Plot early close funding against total amount raised
```{r}

candidateDataEarlyCloseClean %>%
  filter(win == 1) %>%
  ggplot(mapping = aes(x = earlyCloseFundedRelative, y = totalRelative, size = earlyCloseFundedAmt)) + 
  geom_point() + 
  ggtitle("Early Close Funding vs Total Amount Raised") +
  xlab("Relative Early Close Funding") + 
  ylab("Relative Total Raised") +
  labs(
    size = "Early Close Funding ($)"
  ) +
  geom_smooth()

candidateDataEarlyCloseNaN %>%
  filter(win == 1, is.finite(earlyCloseFundedRelative)) %>%
  ggplot(mapping = aes(x = earlyCloseFundedRelative, y = totalRelative, size = earlyCloseFundedAmt)) + 
  geom_point() + 
  ggtitle("Early Close Funding vs Total Amount Raised") +
  xlab("Relative Early Close Funding") + 
  ylab("Relative Total Raised") +
  labs(
    size = "Early Close Funding ($)"
  ) +
  geom_smooth()

```

### Compare early self funding to margin of victory
```{r}
candidateDataEarlyClean %>%
  #filter(earlySelfFundedAmt > 0) %>%
  filter(win == 1) %>%
  ggplot(mapping = aes(x = earlySelfFundedRelative, y = margin)) + 
  geom_point(size=3) + 
  geom_smooth()
```
