---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the janitor and tidyverse packages
```{r}
#install.packages("janitor")
library(janitor)
#install.packages("tidyverse")
library(tidyverse)
library(lubridate)
```

### Get campaign finance data
#### This data includes 2020 Q1 and 2019 semi-annual reports
```{r}

url1 <- "https://cf.ncsbe.gov/CFOrgLkup/ExportDetailResults/?ReportID=173751&Type=REC&Title=COMMITTEE%20TO%20ELECT%20KIM%20COLEY%20-%202020%20First%20Quarter"

url2 <- "https://cf.ncsbe.gov/CFOrgLkup/ExportDetailResults/?ReportID=170343&Type=REC&Title=COMMITTEE%20TO%20ELECT%20KIM%20COLEY%20-%202019%20Organizational"

url3 <- "https://cf.ncsbe.gov/CFOrgLkup/ExportDetailResults/?ReportID=171712&Type=REC&Title=COMMITTEE%20TO%20ELECT%20KIM%20COLEY%20-%202019%20Year%20End%20Semi-Annual"

coley_rcpts_2020_q1 <- read_csv(url1, 
    col_types = cols(
      `Account Abbr` = col_character(), 
        City = col_character(), 
      `Country Name` = col_character(), 
        Date = col_date(format = "%m/%d/%Y"), 
        Description = col_character(), 
      `Employers Name` = col_character(), 
        `Full Zip` = col_character(), 
      `Outside US Postal Code` = col_character(), 
        Profession = col_character(), 
      Purpose = col_character(), 
        State = col_character(), 
      `Street 1` = col_character(), 
        `Street 2` = col_character()),
    skip = 1)

coley_rcpts_2019_1 <- read_csv(url2, 
    col_types = cols(
      `Account Abbr` = col_character(), 
        City = col_character(), 
      `Country Name` = col_character(), 
        Date = col_date(format = "%m/%d/%Y"), 
        Description = col_character(), 
      `Employers Name` = col_character(), 
        `Full Zip` = col_character(), 
      `Outside US Postal Code` = col_character(), 
        Profession = col_character(), 
      Purpose = col_character(), 
        State = col_character(), 
      `Street 1` = col_character(), 
        `Street 2` = col_character()),
    skip = 1)

coley_rcpts_2019_2 <- read_csv(url3, 
    col_types = cols(
      `Account Abbr` = col_character(), 
        City = col_character(), 
      `Country Name` = col_character(), 
        Date = col_date(format = "%m/%d/%Y"), 
        Description = col_character(), 
      `Employers Name` = col_character(), 
        `Full Zip` = col_character(), 
      `Outside US Postal Code` = col_character(), 
        Profession = col_character(), 
      Purpose = col_character(), 
        State = col_character(), 
      `Street 1` = col_character(), 
        `Street 2` = col_character()),
    skip = 1)

# Combine all finance reports together
coley_rcpts_2020 = union_all(coley_rcpts_2020_q1, coley_rcpts_2019_1, coley_rcpts_2019_2)

#Finally, we should rename the columns to remove spaces and generally promote brevity.
names(coley_rcpts_2020) <- c("date","prior","donor","street1","street2","city","state","zip","country","postal","profession","employer","purpose","type","account","payment_form","description","amount","sum_to_date")

```

### Stats grouped by month
```{r}
coley_rcpts_2020_by_month <- coley_rcpts_2020 %>%
  # uncomment the line below to filter out Individual Aggregated Contribution donors
  #filter(donor != "Individual Aggregated Contribution") %>%
  # group by months here, use the month label string instead of numbers, and don't abbreviate
  group_by(Month=month(date, label=TRUE, abbr=FALSE)) %>%
  # summarise our data here
  summarise(
    # sum the dollar amount of donations and round to nearest cent
    TotalAmount = format(round(sum(amount), 2), nsmall = 2), 
    # get the number of distinct donors
    NumberOfDonors = n_distinct(donor),
    # calculate the mean dollar amount and round to nearest cent
    AverageDonation = format(round(mean(amount), 2), nsmall = 2), 
    # calculate the median dollar amount and round to nearest cent
    MedianDonation = format(round(median(amount), 2), nsmall = 2), 
    # calculate the percent of total dollar amount that each month makes up
    PercentOfTotal = sum(amount)/sum(coley_rcpts_2020$amount)
  ) %>%
  # make the month values into the row headers
  remove_rownames %>% column_to_rownames(var="Month") %>%
  # rotate the dataframe by turning it into a matrix, transposing, and converting back into a dataframe
  as.matrix() %>% t() %>% as.data.frame()

coley_rcpts_2020_by_month
```

### Stats grouped by weekday
```{r}
coley_rcpts_2020_by_weekday <- coley_rcpts_2020 %>%
  # uncomment the line below to filter out Individual Aggregated Contribution donors
  #filter(donor != "Individual Aggregated Contribution") %>%
  # group by weekday here
  group_by(Weekday=weekdays(date)) %>%
  # summarise our data here
  summarise(
    # sum the dollar amount of donations and round to nearest cent
    TotalAmount = format(round(sum(amount), 2), nsmall = 2), 
    # get the number of distinct donors
    NumberOfDonors = n_distinct(donor),
    # calculate the mean dollar amount and round to nearest cent
    AverageDonation = format(round(mean(amount), 2), nsmall = 2), 
    # calculate the median dollar amount and round to nearest cent
    MedianDonation = format(round(median(amount), 2), nsmall = 2), 
    # calculate the percent of total dollar amount that each weekday makes up
    PercentOfTotal = sum(amount)/sum(coley_rcpts_2020$amount)
  ) %>%
  # make the weekday values into the row headers
  remove_rownames %>% column_to_rownames(var="Weekday") %>%
  # rotate the dataframe by turning it into a matrix, transposing, and converting back into a dataframe
  as.matrix() %>% t() %>% as.data.frame()

coley_rcpts_2020_by_weekday
```

### Stats grouped by zipcode
```{r}
census_data <- read_csv("nc-zip-code-population.csv",
                        col_types = cols(zip = col_character()))

# join receipts with census data
coley_rcpts_2020_with_income <- left_join(coley_rcpts_2020, census_data)

coley_rcpts_2020_by_zipcode <- coley_rcpts_2020_with_income %>%
  # uncomment the line below to filter out Individual Aggregated Contribution donors
  #filter(donor != "Individual Aggregated Contribution") %>%
  mutate(zip = str_sub(zip, 1, 5)) %>%
  # replace N/A values of zip with "Unknown"
  replace_na(list(zip="Unknown")) %>%
  # group by weekday here
  group_by(Zipcode=zip) %>%
  # summarise our data here
  summarise(
    # sum the dollar amount of donations and round to nearest cent
    TotalAmount = format(round(sum(amount), 2), nsmall = 2), 
    # get the number of distinct donors
    NumberOfDonors = n_distinct(donor),
    # calculate the mean dollar amount and round to nearest cent
    AverageDonation = format(round(mean(amount), 2), nsmall = 2), 
    # calculate the median dollar amount and round to nearest cent
    MedianDonation = format(round(median(amount), 2), nsmall = 2), 
    # calculate the percent of total dollar amount that each weekday makes up
    PercentOfTotal = sum(amount)/sum(coley_rcpts_2020$amount),
    # calculate the donors per capita for Wake county
    DonorsPerCapita = n_distinct(donor)/ first(population),
    # calculate the dollar amount per capita for Wake county,
    DollarsPerCapita = sum(amount)/ first(population),
    # calculate the median donor amount over the median income
    MedianDonationOverIncome = median(amount)/ first(median_income)
  ) %>%
  # make the weekday values into the row headers
  remove_rownames %>% column_to_rownames(var="Zipcode") %>%
  # rotate the dataframe by turning it into a matrix, transposing, and converting back into a dataframe
  as.matrix() %>% t() %>% as.data.frame()

coley_rcpts_2020_by_zipcode
```

### Reporting Question
#### I noticed that a large amount of donations came from Kim Coley herself as in-kind donations. I'm curious about how well campaigns turn out depending on how much the candidate invests in themselves. Is there a correlation between how much money a candidate invests in their own campaign and how much money others invest in their campaign? Do candidates who invest a certain amount into their campaign have a higher chance of winning?

### Visualization 1
#### Here we're visualizing the difference in total amount donated based on payment type.
```{r}
coley_rcpts_2020 %>%
  group_by(payment_form) %>%
  summarise(TotalAmount = round(sum(amount), 2)) %>%
  # plot total amount by payment form
  ggplot(mapping = aes(x = payment_form, y = TotalAmount)) +
  geom_bar(stat = "identity") +
  # label dollar amounts for each bar
  geom_text(aes(label = format(round(TotalAmount, 2), nsmall=2)), vjust = -0.3, size = 3.5)+
  theme_minimal()
```

### Visualization 2
#### Here we're visualizing the payment amounts against the median salary for the donor's zipcode
```{r}
# we define this function for formating axis labels to whole dollar amounts
scaleFUN <- function(x) sprintf("%.0f", x)

coley_rcpts_2020_with_income %>%
  # remove rows without median income values available
  drop_na(median_income) %>%
  group_by(median_income) %>%
  ggplot(mapping = aes(x = median_income, y = amount)) + 
  geom_point(size=3) + 
  geom_smooth() +
  # format axis labels to whole dollar amounts
  scale_y_continuous(labels=scaleFUN) +
  scale_x_continuous(labels=scaleFUN)
```
  
  