---
title: "Class Notes Feb. 5, 2020"
output: html_notebook
---

Learning R is like learning a new spoken language. We just need to learn the words and rules.

These are the three key verbs/functions we'll be using this semester:

- `arrange`
- `filter`
- `select`

Journalists want to find things that are unexpected (not normal). All three of these functions help us do that. `arrange` shows us the extremes of our data, `filter` shows us rows that meet certain criters, and `select` shows us only certain columns.

We'll be using the same MLBpayrolls dataframe for the next few assignments.

## Learning arrange

The `arrange` function sorts our dataframe (ascending or descending) by specific columns.

`arrange` takes the following arguments:

1. The dataframe we want to arrange
2. The column we want to sort by
    - To sort in descending order, you must wrap the column name in `desc()`
3. (optional) Any additional columns we want to sort by
    - Additional sort columns are used to break ties in previous columns
    
```{r}
# first we need to load tidyverse to gain access to the arrange function
library(tidyverse)

# here we arrange MLBpayrolls by the 2016 payroll in ascending order, and break ties with 2004 payroll in descending order
arrange(MLBpayrolls, Payroll.2016, desc(Payroll.2004))
```

Tips:

- You need to load tidyverse only once everytime you reload the RStudio application
- If there are spaces in your arguments, you need to wrap them in tick marks ( \` ).
  - Eg. \`Payroll 2016\`
- If you use single quote marks ( ' ) instead of tick marks, you may get an "unexpected string" error
- If you have Payroll.2016 (with a dot) that's ok
  - At the top of HW6 when we load the Excel spreadsheet, the `.name_repair = "universal"` argument replaces all the spaces with dots


## Learning filter

`filter` gives us only the rows that meet certain criteria.

`filter` takes the following arguments:

1. The dataframe we want to filter
2. The criteria we want to filter by
3. (optional) Any additional criteria we want to filter by

```{r}
# here we filter MLBpayrolls to show only teams (rows) that paid more than 150,000,000 in 2016
filter(MLBpayrolls, Payroll.2016 > 150000000)

# here we filter MLBpayrolls to show only teams that paid more in 2016 than the average payroll in 2016
filter(MLBpayrolls, Payroll.2016 > mean(MLBpayrolls$Payroll.2016))

# here we filter MLBpayrolls to show only teams that paid more in 2016 than the average payroll in 2016 AND paid more in 2010 than the average payroll in 2016
filter(MLBpayrolls, Payroll.2016 > mean(MLBpayrolls$Payroll.2016), Payroll.2010 > mean(MLBpayrolls$Payroll.2016))

# here we filter MLBpayrolls to show only teams that paid more in 2016 than the average payroll in 2016 OR paid more in 2010 than the average payroll in 2016
filter(MLBpayrolls, Payroll.2016 > mean(MLBpayrolls$Payroll.2016) | Payroll.2010 > mean(MLBpayrolls$Payroll.2016))
```

Tips:

1. Separating criteria by commas means "AND"
2. Separating criteria by pipes ( | ) means "OR"
3. You can see logical operators here: http://learn.r-journalism.com/en/wrangling/dplyr/dplyr/#logical-operators

### Eg. Get all the information for just the teams from New York.

```{r}
# here we filter MLBpayrolls to show only teams that are called "New York Mets" and "New York Yankees"
filter(MLBpayrolls, TEAM %in% c("New York Mets","New York Yankees"))
# same as above
filter(MLBpayrolls, TEAM == "New York Mets" | TEAM == "New York Yankees")
# same as above
NYTeams <- c("New York Mets","New York Yankees")
filter(MLBpayrolls, TEAM %in% NYTeams)
```

When we run multiple commands in a single chunk (like above), only the results of last one will be shown in the R Markdown file. Make sure that the last command gives the results that you want others to see.

## Learning select

The `select` function gives us the only specific columns.

We can use a pipe `%>%` to send results from one command to another. The pipe goes between commands and says "and then do this".

```{r}
# here we take MLBpayrolls 
# then we filter it to show only teams that paid more in 2016 than the average payroll in 2016
# then we take our filter result and select only the TEAM and Payroll 2016 columns to be shown
# then we take our select result and arrange it in descending order by Payroll 2016
MLBpayrolls %>%
filter(Payroll.2016 > mean(MLBpayrolls$Payroll.2016)) %>%
select(TEAM, Payroll.2016) %>%
arrange(desc(Payroll.2016))
```

Tips:

- `select` removes columns, so the order of your command pipeline matters
  - If you `select` to remove a column, you won't be able to use `arrange` and sort by that removed column later
  
Feedback on readings and HW:
- Students think it will be more useful to have R readings, then go over R in class, and then do homework using what we've learned
- Ryan doesn't want us to spend more than 3 hours on a homework assignment

Homework Info:
- In the `read_excel` function at the top of the homework, the first argument is the path to the Excel file. The `..` at the beginning of the path tells R to look up one directory/folder in our file system.
  - If you don't have a `data` folder in the outside of your R folder, you need to change this path
- In the `read_excel` function at the top of the homework, the `.name_repair = "universal"` argument replaces all the spaces with dots